---
title: "Analyzing Palmer Penguins with R"
author: "Delaney Woods"
date: "2026-09-22"
---

## Introduction

In this post, I use R to explore the Palmer Penguins dataset. I look at body mass across penguin species and examine the relationship between flipper length and body mass.

The data comes from the [Palmer Penguins project](https://allisonhorst.github.io/palmerpenguins/), which uses data from the Palmer Station Antarctica LTER.

## Loading the data

```{r}
#| message: false
#| warning: false

library(palmerpenguins)

penguins <- palmerpenguins::penguins

head(penguins)
```
## Body mass by species

I calculated the average body mass for each penguin species to compare their typical body sizes.

```{r}
avg_species_mass <- aggregate(body_mass_g ~ species, data = penguins, FUN = mean, na.rm = TRUE)

avg_species_mass
```

## Flipper length and body mass

This figure shows the relationship between flipper length and body mass for the three penguin species.

```{r}
#| label: fig-flipper-mass-r
#| fig-cap: "Flipper length and body mass of Palmer Penguins, by species."


library(ggplot2)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
geom_point() +
scale_color_manual(values = c(
"Adelie" = "#A8DADC",
"Chinstrap" = "#F4A6A6",
"Gentoo" = "#CDB4DB"
)) +
labs(
x = "Flipper length (mm)",
y = "Body mass (g)",
color = "Species",
shape = "Species"
)
```

