---
title: "Analyzing Palmer Penguins Using R"
author: "Delaney Woods"
date: "2026-09-22"
---

## Introduction

In this post, I use R to explore the Palmer Penguins dataset. I look at body mass across penguin species and examine the relationship between flipper length and body mass.

The data comes from the [Palmer Penguins project](https://allisonhorst.github.io/palmerpenguins/), which uses data from the Palmer Station Antarctica LTER.

## Loading the data

```{r}
#| message: false
#| warning: false

library(palmerpenguins)

penguins <- palmerpenguins::penguins

head(penguins)
```
## Body mass by species

I calculated the average body mass for each penguin species to compare their typical body sizes.

```{r}
avg_species_mass <- aggregate(body_mass_g ~ species, data = penguins, FUN = mean, na.rm = TRUE)

avg_species_mass

```

## Summary statistics

```{r}

summary_by_species <- aggregate(
  cbind(flipper_length_mm, body_mass_g) ~ species,
  data = penguins,
  FUN = mean,
  na.rm = TRUE
)

summary_by_species

```

## Flipper length and body mass

This figure shows the relationship between flipper length and body mass for the three penguin species.

```{r}
#| label: fig-flipper-mass-r
#| fig-cap: "Flipper length and body mass of Palmer Penguins, by species."


library(ggplot2)

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species, shape = species)) +
geom_point() +
scale_color_manual(values = c(
"Adelie" = "#A8DADC",
"Chinstrap" = "#F4A6A6",
"Gentoo" = "#CDB4DB"
)) +
labs(
x = "Flipper length (mm)",
y = "Body mass (g)",
color = "Species",
shape = "Species"
)
```

## Findings
Based on the graph, it is clear that flipper length tends to increase as body mass increases. In particular, Gentoo are the largest penguin species and have both the heaviest body mass and the longest flipper length.
